---
title: "p8105_hw3_gw2383"
author: "Guojing Wu"
date: "2018/10/7"
output: html_document
# output:
#   github_document:
#     toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
theme_set(theme_bw())
```

## Problem 1

### data cleaning

```{r}
brfss_df =
  p8105.datasets::brfss_smart2010 %>%
  janitor::clean_names() %>% # cleaning
  rename(location_abbr = locationabbr, location_desc = locationdesc) %>% # format the variable name
  filter(topic == "Overall Health") %>% # focus on the “Overall Health” topic
  filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>% # choose response
  arrange(response) # organize responses
```

### some questions

* In 2002, `r names(which(brfss_df %>% filter(year == 2002) %>% distinct(location_abbr, location_desc) %>% select(location_abbr) %>% table() == 7))` states were observed at 7 locations

* spaghetti plot

```{r}
brfss_df %>%
  distinct(year, location_abbr, location_desc) %>%
  group_by(year, location_abbr) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = year, y = n, group = location_abbr, color = location_abbr)) +
  geom_line() +
  labs(
    y = "number of observatons",
    title = " number of observations in each state from 2002 to 2010"
  ) +
  theme_bw()
```

* a table of mean and standard deviation

```{r}
brfss_df %>%
  filter(year %in% c(2002, 2006, 2010) & location_abbr == "NY" & response == "Excellent") %>%
  group_by(year) %>%
  summarise(mean_exc = mean(data_value),
            sd_exc = sd(data_value)) %>% 
   knitr::kable(digits = 1)
```

* a five-panel plot

```{r}
brfss_df %>%
  group_by(year, location_abbr, response) %>%
  summarise(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, group = location_abbr, color = location_abbr)) +
  geom_line() +
  labs(y = "average proportion") +
  facet_grid(response ~ ., scales = "free") + # ploting vertically and without the same axis scaling
  theme_bw()
```

## Problem 2

```{r}
insta_df = p8105.datasets::instacart %>% 
  janitor::clean_names()
```

### short description

There are `r dim(insta_df)[2]` variables and `r dim(insta_df)[1]` observations in this dataset. It contains details of every order: like `user id`: customer identifier, `order id`: order identifier, `product id`: product identifier, `product name`, `reordered`: whether it's reordered or not, when it was ordered (assume 0 was assigned to Sunday). Also, in the local store, there are `r dim(distinct(insta_df, department))[1]` departments.

```{r} 
insta_df %>% select(user_id, order_id, product_id, product_name, reordered, order_dow) %>% head(10) %>% knitr::kable()
```

### some questions

* There are `r dim(distinct(insta_df, aisle))[1]` aisles in the store. `r (insta_df %>% group_by(aisle) %>% summarise(item_sum = n()) %>% arrange(desc(item_sum)))[1,1]` aisle is the most items ordered from.

* a plot that shows the number of items ordered in each aisle

```{r, fig.height = 20}
# a function for sorting the factors so that aisles will be listed in descending way
reorder_size <- function(x) {
  factor(x, levels = names(sort(table(x))))
}

insta_df %>% 
  select(aisle) %>% 
  ggplot(aes(x = reorder_size(aisle))) +
  geom_bar() +
  scale_y_log10() +
  labs(x = "aisles", 
       y = "log10(total number of items)", 
       title = "number of items ordered in each aisle") +
  coord_flip()
```

* the most popular item:

```{r}
insta_df %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(item_num = n()) %>% 
  group_by(aisle) %>% 
  mutate(item_rank = min_rank(desc(item_num))) %>% # calculate the rank
  filter(item_rank == 1) %>% 
  knitr::kable(digits = 1)
```

* a table showing the mean hour of the day

for the consecutive number in order_dow, we assume 0 assign to Sunday, 1 assign to Monday and so on ...

```{r}
insta_df %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour) %>% 
  rename(Sunday = "0", Monday = "1", Tuesday = "2", Wednesday = "3", Thursday = "4", Friday = "5", Saturday = "6") %>% # rename it to make it readable
  knitr::kable(digit = 1)
```

## Problem 3

```{r}
noaa_df = p8105.datasets::ny_noaa %>% 
  janitor::clean_names()
```

### short description

There are `r dim(noaa_df)[2]` variables and `r dim(noaa_df)[1]` observations in this dataset. It contains details of everyday weatherr: like `id`: Weather station ID, `date`: Date of observation, `prcp`: Precipitation (tenths of mm), `snow`: Snowfall (mm), `snwd`: Snow depth (mm), `tmax`: aximum temperature (tenths of degrees C), `tmin`: Minimum temperature (tenths of degrees C).

Also we noticed that there are many NAs in the dataset, below are the table of the proportion of NAs in each column

```{r}
noaa_df %>% 
  summarise(id_na = sum(is.na(id))/length(id),
            date_na = sum(is.na(date))/length(date),
            prcp_na = sum(is.na(prcp))/length(prcp),
            snow_na = sum(is.na(snow))/length(snow),
            snwd_na = sum(is.na(snwd))/length(snwd),
            tmax_na = sum(is.na(tmax))/length(tmax),
            tmin_na = sum(is.na(tmin))/length(tmin)
            ) %>% 
  knitr::kable(digits = 3)
```

### some questions

* data cleaning and create separate variables for year, month, and day

```{r}

```

